---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# visitalaneysluverds

<!-- badges: start -->
<!-- badges: end -->


The package provides an easy way to fetch the CPI from Statistics Iceland.

## Installation

You can install the development version of visitalaneysluverds from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("bgautijonsson/visitalaneysluverds")
```

## Example

```{r example}
library(visitalaneysluverds)
d <- vnv()

d
```


```{r}
plot(cpi ~ date, data = d, type = "l")
```

You can input a date on which the CPI should be equal to 1. The date has to be the first day of a month between 1988-05-01 and the month preceding the current month. If no date is input the CPI is equal to 1 ni 1988-05-01

```{r}
d <- vnv(date_unity = as.Date("2018-01-1"))
plot(cpi ~ date, data = d, type = "l")
```

