---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# visitalaneysluverds

<!-- badges: start -->
<!-- badges: end -->


The package provides an easy way to fetch the CPI from Statistics Iceland.

## Installation

You can install the development version of visitalaneysluverds from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("bgautijonsson/visitalaneysluverds")
```

## Example

### CPI table

#### Monthly CPI

```{r}
library(visitalaneysluverds)
d <- vnv()

d
```


```{r}
plot(cpi ~ date, data = d, type = "l")
```

You can input a date on which the CPI should be equal to 1. The date has to be between 1988-05-01 and the month preceding the current month. If no date is input the CPI will be equal to 1 in the previous month.

```{r}
d <- vnv(date_unity = as.Date("2018-01-1"))
plot(cpi ~ date, data = d, type = "l")
```

#### Yearly CPI

```{r}
d <- vnv_yearly()

d
```


```{r}
plot(cpi ~ year, data = d, type = "l")
```

You can input a year on which the CPI should be equal to 1. The year has to be between 1988 and the current year. If no year is input the CPI will be equal to 1 in the current year.

```{r}
d <- vnv_yearly(year_unity = 2002)
plot(cpi ~ year, data = d, type = "l")
```

### Performing CPI adjustments

#### Monthly prices

The package offers use of the function `vnv_convert()` to convert prices to a common CPI standard.

```{r}
price <- c(1, 2, 3, 4, 5)
date <- as.Date(c("2020-01-01", "2019-01-01", "2018-01-01", "2017-01-01", "2016-01-01"))
d <- data.frame(
    date = date,
    price = price
)

d$price_2020_01_01 <- vnv_convert(d$price, d$date, date_unity = as.Date("2020-01-01"))

d
```


The function `vnv_convert()` works because the package downloads CPI data from Statistics Iceland once per session and makes it available to internal functions, allowing the user to use the most recent CPI data to compare prices if needed.

#### Yearly prices

The `vnv_convert()` function detects whether the `obs_date` and `date_unity` variables have the class `Date` or `integer` and automatically performs the right type of CPI adjustmeny (monthly/yearly) for your data.

```{r}
price <- c(1, 2, 3, 4, 5)
year <- c(2011, 2012, 2013, 2014, 2015)
d <- data.frame(
    year = year,
    price = price
)

d$price_2020 <- vnv_convert(d$price, d$year, date_unity = 2020)

d
```

